---
title: "PROJECT"
Author: Ojaswi Gupta
output: html_notebook
---

```{r}
library(data.table)
library(dplyr)
library(ggplot2)
library(tidyr)
library(tidyverse)
library(quantmod)
library(TTR)
library(plotly)
library(PerformanceAnalytics)
```

```{r}
origwd <- getwd()
print(origwd)

start <- Sys.time()
corepath <-"/Users/ojaswigupta/Desktop/coursework/sem2/Inro to DS with R/Intro to DS with R Project/"
setwd(corepath)
```

loading the dataset with ESG data. change the path to 
```{r}
load(paste0(corepath,"data_esg.RData"))
```

There are 6 attributes:      
- closing price (**Close**),     
- past (realized) 1 month volatility (**Vol_1M**),     
- market capitalisation in M$ (**Mkt_Cap**),     
- price-to-book ratio (**P2B**),     
- debt-to-equity (**D2E**),    
- ESG ranking (**ESG_rank** from Sustainalytics, in %).

To modify the dataset, separating individual values for each stock, with all stocks aligned horizontally.

```{r}
Column_headers<-colnames(data_esg)
attribute_list<- Column_headers[3:8]
attribute_list
```

Convert the data from stacked on top of each other to horizontally aligned by Tick for every value type.

```{r}
for (i in attribute_list){
    df<-pivot_wider(data_esg[,c("Tick","Date",i)],names_from="Tick", values_from=i)
    assign(paste('data_',i,sep=''),df)
}
Tick<-unique(data_esg$Tick)
```

Filter out the stocks with a high ESG rating. The threshold here is 75.
Filter by ESG and use it.
```{r}
ESG_signal<-c() #call it signal 
ESG_signal <- data_ESG_rank
stock_filter1<-c()
stock_filter1 <- data_Close

for (i in 1:nrow(data_ESG_rank)){
  for (j in 2:ncol(data_ESG_rank)){
    if (data_ESG_rank [i,j]>75){
    ESG_signal[i,j]<-1
    }
    else {ESG_signal[i,j]<-0}
    stock_filter1[i,j]<-ESG_signal[i,j]
    }
}
tail(stock_filter1)
```

EMA 7 months and 2 months to mimic the 150 day and 50 day moving average. 
```{r}
Mov_Avg1<-c()
Mov_Avg2<-c()
for (i in Tick){
  EMA_7<-data.table(EMA(data_Close[i],n=7))
  EMA_7<-bind_cols(EMA_7)
  
  Mov_Avg1<-cbind(Mov_Avg1,EMA_7)
  
  EMA_2<-data.table(EMA(data_Close[i],n=2))
  EMA_2<-bind_cols(EMA_2)
  
  Mov_Avg2<-cbind(Mov_Avg2,EMA_2)
}
colnames(Mov_Avg1)=Tick
Mov_Avg1<-cbind(data_Close$Date,Mov_Avg1)
colnames(Mov_Avg1)[1]="Date"

colnames(Mov_Avg2)=Tick
Mov_Avg2<-cbind(data_Close$Date,Mov_Avg2)
colnames(Mov_Avg2)[1]="Date"

EMA_diff<-Mov_Avg2[,2:ncol(Mov_Avg2)]-Mov_Avg1[,2:ncol(Mov_Avg1)]
EMA_diff<-cbind(data_Close$Date,EMA_diff)
colnames(EMA_diff)[1]="Date"
EMA_diff[is.na(EMA_diff)]<-0

class(EMA_diff)<-class(as.data.frame(EMA_diff))
```

Using the EMA to create signal. Whenever, the 2 month EMA crosses the 7 month EMA we set thesigna l as 1 (Buy). When they are equal or 7 month EMA is greater, we set the signal at 0 (Sell)
```{r}
EMA_signal<-c()
EMA_signal <- EMA_diff
stock_filter2<-c()
stock_filter2 <- data_Close
for (i in 1:nrow(EMA_diff)){
  for (j in 2:ncol(EMA_diff)){
    if (EMA_diff [i,j]<0){
    EMA_signal[i,j]<-0
    }
    else {EMA_signal[i,j]<-1}
    stock_filter2[i,j]<- EMA_signal[i,j]
    }
}
tail(stock_filter2)
```

Rate of change for stocks to be calculated. using rate of change as a sell signal with the condition of booking profits when it is more than 15%. This is and attempt to eliminate the high volatility stocks.

```{r}
rate_of_change<-c()
for (i in Tick){
 roc<-data.table(ROC(data_Close[i],n=2))
  roc<-bind_cols(roc)
  
  rate_of_change<-cbind(rate_of_change,roc)
}
colnames(rate_of_change)=Tick
rate_of_change<-cbind(data_Close$Date,rate_of_change)
colnames(rate_of_change)[1]="Date"
rate_of_change[is.na(rate_of_change)]<-0
class(rate_of_change)<-class(as.data.frame(rate_of_change))

roc_signal<-c()
roc_signal <- EMA_diff
stock_filter3<-c()
stock_filter3 <- data_Close
for (i in 1:nrow(rate_of_change)){
  for (j in 2:ncol(rate_of_change)){
    if (rate_of_change [i,j]>.15 & rate_of_change [i,j]< -.15){
    roc_signal[i,j]<-0
    }
    else {roc_signal[i,j]<-1}
    stock_filter3[i,j]<-roc_signal[i,j]
    }
}
tail(stock_filter3)
```

Using RSI as an indicator for selling off the stocks continuously underperforming the market. Whenever RSI is below 

```{r}
RSI<-c()

for (i in Tick){
  rsi<- data.table(RSI(data_Close[i]))
  RSI<-cbind(RSI,rsi)
}
colnames(RSI)=Tick
RSI<-cbind(data_Close$Date,RSI)
colnames(RSI)[1]="Date"

RSI[is.na(RSI)]<-50

class(RSI)<-class(as.data.frame(RSI))

RSI_filter<-c()
RSI_filter <- RSI
stock_filter4<-c()
stock_filter4 <- data_Close

for (i in 1:nrow(RSI)){
  for (j in 2:ncol(RSI)){
    if (RSI [i,j]<35 & RSI [i,j]>85){
    RSI_filter[i,j]<-0
    }
    else {RSI_filter[i,j]<-1}
    stock_filter4[i,j]<- RSI_filter[i,j]
    }
}

tail(stock_filter4)
```


```{r}
Bollinger_band<-data.frame()
n<-1
for (i in Tick){
  bband<- data.table(BBands(data_Close[i]),n=10,sd=2)
  if(n==1){
    Bollinger_band<-bband$pctB
  }
  else{
  Bollinger_band<-cbind(Bollinger_band,bband$pctB)  
  }
  n<-2
}
Bollinger_band <-as.data.frame(Bollinger_band)
colnames(Bollinger_band)=Tick
Bollinger_band<-cbind(data_Close$Date,Bollinger_band)
colnames(Bollinger_band)[1]="Date"

Bollinger_band[is.na(Bollinger_band)]<-0

Bollinger_filter<-c()
Bollinger_filter<- RSI_filter
stock_filter5<-c()
stock_filter5 <- data_Close

class(Bollinger_filter)<-class(as.data.frame(Bollinger_filter))

for (i in 1:nrow(Bollinger_band)){
  for (j in 2:ncol(Bollinger_band)){
    if (Bollinger_band [i,j]>1){
    Bollinger_filter[i,j]<-0
    }
    else {Bollinger_filter[i,j]<-1}
    stock_filter5[i,j]<- Bollinger_filter[i,j]
    }
}
tail(stock_filter5)
```

Create a Trade signal using all the filters estimated above.

```{r}
Trade_signal<-data_Close[,2:ncol(data_Close)]*
stock_filter1[,2:ncol(stock_filter1)]*
stock_filter2[,2:ncol(stock_filter2)]*
stock_filter3[,2:ncol(stock_filter3)]*
stock_filter4[,2:ncol(stock_filter4)]*
stock_filter5[,2:ncol(stock_filter5)]

Trade_signal<-cbind(data_Close$Date,Trade_signal)
colnames(Trade_signal)[1]="Date"
tail(Trade_signal)
```


This part of code enables the buyiung and selling of stocks based on the above calculated filters. 
The way we proceed here is as follows:
-Take the initial investment amount and divide it equally among all eligible stocks at T1.
-square off the posiitons at T2 and then again redistribute the new capital among the eligible stocks at T2.
-repeat the process until the end of time series of data.

```{r}
Money<-1000000
Temp<-1
Test<-data.frame()
share_price_quantity<-data.frame("price"=0,"shares"=0,"i"=0,"j"=0)

for (i in 1:nrow(Trade_signal)){
  to_buy<-0
if (Temp!=1){
  for (k in 2:nrow(share_price_quantity)){
      Money<-Money+(data_Close[i,share_price_quantity[k,4]])*share_price_quantity[k,2] 
  }
}
  Test<-append(Test,c(Money))
  share_price_quantity<-data.frame("price"=0,"shares"=0,"i"=0,"j"=0)
  #to count the number of unique shares to be bought
  for (j in 2:ncol(Trade_signal)){
    if (Trade_signal[i,j]!=0){
      to_buy<-to_buy+1
    }
  }
  if(to_buy==0){next}
#money available for each unique company share to buy them
Moneyps<-Money/to_buy
  for (j in 2:ncol(Trade_signal)){
    if (Trade_signal[i,j]!=0){
      price<-Trade_signal[i,j]
      shares<-Moneyps/Trade_signal[i,j]
      share_price_quantity[nrow(share_price_quantity)+1,]=c(price,shares,i,j)
      Money<-Money-shares*price
}
}
  Temp<-2
}
```

```{r}
Final_result<-c()
Final_result<-do.call(rbind.data.frame, Test)
Final_result<-cbind(data_Close$Date,Final_result)
colnames(Final_result)[1]="Date"
colnames(Final_result)[2]="Money_pf"
Final_result<-cbind(Final_result,100*ROC(Final_result$Money_pf,n=1,type="discrete"))
colnames(Final_result)[3]="Return_pf"
Final_result<-drop_na(Final_result)
tail(Final_result)
```

Summary statistics for total data without any filters and with an equally weighted portfolio.
Calculating the monthly returns and overall growth of the investment.

```{r}
Money_pf<-1000000  #million dollars to begin with
Money_ewpf<-Money_pf/(ncol(data_Close)-1)
weight<-c()
portfolio_ew<-c()
for (i in 2:ncol(data_Close)){
  w<-Money_ewpf/data_Close[1,i]
  weight<-bind_cols(weight,w)
}
colnames(weight)=Tick
weight<-cbind(data_Close$Date,weight)
colnames(weight)[1]="Date"

portfolio_ew<-weight[,2:383]*data_Close[,2:383]
total<-as.data.frame(rowSums(portfolio_ew[,1:382]))
total<-cbind(data_Close$Date,total)
colnames(total)[1]="Date"
colnames(total)[2]="Money_ew"
total<-cbind(total,100*ROC(total$Money_ew,n=1,type="discrete"))
colnames(total)[3]="Return_ew"
total<-drop_na(total)
```

Comparison of equally weighted and filtered portfolios.

```{r}
Comparison<-data.frame()
Comparison<-total
Comparison<-cbind(Comparison,Final_result[,2:3])
tail(Comparison)
```


```{r}
plot1<-ggplot(Comparison, aes(x=Date)) + 
  geom_line(aes(y = Money_ew), color = "maroon", linetype="dotted") + 
  geom_line(aes(y = Money_pf), color="steelblue") +
  ylab("Total Investment")
# Turn it interactive with ggplotly
plot1 <- ggplotly(plot1)
plot1
```

```{r}
plot2<-ggplot(Comparison, aes(x=Date)) + 
  geom_line(aes(y = Return_ew), color = "maroon", linetype="dotted") + 
  geom_line(aes(y = Return_pf), color="steelblue") +
  ylab("Monthly Returns (%)")
# Turn it interactive with ggplotly
plot2 <- ggplotly(plot2)
plot2
```

Comparison statistics:
```{r}
vol_ew <- sd(Comparison$Return_ew) * sqrt(12)
vol_pf <- sd(Comparison$Return_pf) * sqrt(12)
StDev_ew<- sd(Comparison$Return_ew)
StDev_pf<- sd(Comparison$Return_pf)

returns_ew<-xts(Comparison$Return_ew,order.by=as.Date(Comparison$Date))
colnames(returns_ew)<-"returns_ew"
returns_pf<-xts(Comparison$Return_pf,order.by=as.Date(Comparison$Date))
colnames(returns_pf)<-"returns_pf"

Sharpe_ew<-SharpeRatio(returns_ew,Rf=0.002,FUN='StdDev')*sqrt(12)
Sharpe_pf<-SharpeRatio(returns_pf,Rf=0.002,FUN='StdDev')*sqrt(12)

Results<- data.frame("Portfolio" = c("Benchmark","Managed"), "Sharpe_Ratio" = c(Sharpe_ew,Sharpe_pf), "St_Dev" = c(StDev_ew,StDev_pf),"Volatility" = c(vol_ew,vol_pf),"net_percent_Ret" = c(100*((Comparison[84,2]/1000000)-1),100*((Comparison[84,4]/1000000)-1)))
print(Results)

```


```{r}
end <- Sys.time()
runtime<- end-start
runtime
```